Fast Fujisaki-Okamoto transformation using encrypt-then-mac and applications to Kyber

### 2024-08-02 etm-transform/manuscript/main.tex
update

% Custom commands
\newcommand{\monospace}{\texttt}
\newcommand{\pke}{\monospace{PKE}}
\newcommand{\keygen}{\monospace{Gen}}
\newcommand{\encrypt}{\monospace{Enc}}
\newcommand{\decrypt}{\monospace{Dec}}
\newcommand{\kem}{\monospace{KEM}}
\newcommand{\encap}{\monospace{Encap}}
\newcommand{\decap}{\monospace{Decap}}
\newcommand{\etm}{\monospace{EtM}}  % encrypt-then-mac
\newcommand{\mac}{\monospace{MAC}}
\newcommand{\sign}{\monospace{Sign}}
\newcommand{\verify}{\monospace{Verify}}
\newcommand{\pk}{\monospace{pk}}
\newcommand{\sk}{\monospace{sk}}
\newcommand{\pco}{\monospace{PCO}}
\newcommand{\cvo}{\monospace{CVO}}
\newcommand{\leftsample}{\stackrel{\$}{\leftarrow}}
\newcommand{\llbrack}{[\![}
\newcommand{\rrbrack}{]\!]}
\newcommand{\norm}[1]{\left\lvert #1 \right\rvert}
\newcommand{\fotplus}{\monospace{FOT+}}

@@

The problem of ciphertext integrity was solved in symmetric cryptography. Given a semantically secure symmetric cipher and an existentially unforgeable message authentication code, combining them using ``encrypt-then-mac'' provides authenticated encryption \cite{bellare2000authenticated}. We took inspiration from this strategy and applied a similar technique to provide ciphertext integrity for a public-key encryption scheme, which then translates to an \monospace{IND-CCA} secure $\kem$. Using a message authentication code for ciphertext integrity replaces the re-encryption step in decryption with the computation of a tag, which should offer significant performance improvements while maintaining comparable level of security.

The main challenge in applying ``encrypt-then-mac'' to public-key cryptography is the lack of a pre-shared $\mac$ key. We proposed to derive the shared $\mac$ key by hashing the plaintext message. We will prove in section \ref{sec:main-results} that, under the random oracle model, the $\mac$ key is securely hidden behind the hash function, and producing a valid pair of ciphertext and tag without full knowledge of the plaintext constitutes a forgery attack on the message authentication code. Thanks to the modular construction in \cite{hofheinz2017modular}, providing ciphertext integrity in the underlying encryption scheme gives us an \monospace{IND-CCA} secure $\kem$ for free.

@@

\subsubsection{Security}
We discuss the security of a $\pke$ using the sequence of games described in \cite{shoup2004sequences}. Specifically, we first define the $\monospace{OW-ATK}$ and the $\monospace{IND-CPA}$ game as they pertain to a public key encryption scheme. In later section we will define the $\monospace{IND-CCA}$ game as it pertains to a key encapsulation mechanism.

In the \monospace{OW-ATK} game, an adversary's goal is to recover the decryption of a randomly generated ciphertext.

@@

    \begin{algorithm}[H]
        \caption{
            The $\monospace{OW-ATK}$ game
        }\label{alg:pke-ow-atk-game}
        \begin{algorithmic}[1]
            \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
            \State $m^\ast \leftsample \mathcal{M}$
            \State $c^\ast \leftsample \encrypt(\pk, m)^\ast$
            \State $\hat{m} \leftsample \mathcal{A}^{\mathcal{O}_\monospace{ATK}}(1^\lambda, \pk, c^\ast)$
            \State \Return $\llbrack m^\ast = \hat{m} \rrbrack$
        \end{algorithmic}
    \end{algorithm}

    \caption{The \monospace{OW-ATK} game}
    \label{fig:pke-ow-atk-game}
\end{figure}

The adversary $\mathcal{A}$ with access to oracle(s) $\mathcal{O}_\monospace{ATK}$ wins the game if its guess $\hat{m}$ is equal to the challenge plaintext $m^\ast$. The \textit{advantage} $\epsilon_\monospace{OW-ATK}$ of an adversary in this game is the probability that it wins the game.

The choice of oracle(s) $\mathcal{O}_\monospace{ATK}$ depends on the choice of $\monospace{ATK}$. Specifically:

\begin{equation*}
    \mathcal{O}_\monospace{ATK} = \begin{cases}
        - & \monospace{ATK} = \monospace{CPA} \\
        \monospace{PCO} & \monospace{ATK} = \monospace{PCA} \\
        \monospace{CVO} & \monospace{ATK} = \monospace{VA} \\
        \monospace{PCO, CVO} & \monospace{ATK} = \monospace{PCVA}
    \end{cases}
\end{equation*}

Where the definitions of plaintext-checking oracle $\monospace{PCO}$ and the ciphertext-validation oracle $\monospace{CVO}$ are inherited from \cite{hofheinz2017modular}

@@

In the \monospace{IND-CPA} game (algorithm \ref{alg:ind-cpa-game}), an adversary's goal is to distinguish the encryption of one message from the encryption of another message. Given the public key, the adversary outputs two adversarially chosen messages and obtains the encryption of a random choice between these two messages. The adversary wins the \monospace{IND-CPA} game if it correctly identifies which message the encryption is obtained from.

\begin{figure}[H]
    \begin{floatrow}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{The $\monospace{IND-CPA}$ game}\label{alg:ind-cpa-game}

                \begin{algorithmic}[1]
                    \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                    \State $(m_0, m_1) \leftsample \mathcal{A}(a^\lambda, \pk)$
                    \State $b \leftsample \{0, 1\}$
                    \State $c^\ast \leftsample \encrypt(\pk, m_b)$
                    \State $\hat{b} \leftsample \mathcal{A}(1^\lambda, \pk, c^\ast)$
                \State \Return $\llbrack b = \hat{b} \rrbrack$
            \end{algorithmic}
        \end{algorithm}
        }{\caption{The \monospace{IND-CPA} game}}
    \end{floatrow}
\end{figure}

The \textit{advantage} $\epsilon_\monospace{IND-CPA}$ of an \monospace{IND-CPA} adversary $A$ is defined by

$$
    \monospace{Adv}_\monospace{IND-CPA}(A) = \norm{
        P[\hat{b} = b] - \frac{1}{2}
    }
$$

@@

\begin{figure}[H]
    \begin{floatrow}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{
                    \monospace{IND-CCA} game for \kem
                }\label{alg:kem-ind-cca-game}

                \begin{algorithmic}[1]
                    \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                    \State $(c^\ast, k_0) \leftsample \encap(\pk)$
                    \State $k_1 \leftsample \mathcal{K}$
                    \State $b \leftsample \{0, 1\}$
                    \State $\hat{b} \leftsample \mathcal{A}_\monospace{IND-CCA}^{\mathcal{O}_\decap}(
                        1^\lambda, \pk, c^\ast, k_b
                    )$
                    \State \Return $\llbrack \hat{b} = b \rrbrack$
                \end{algorithmic}
            \end{algorithm}
        }{\caption{The \monospace{KEM-IND-CCA2} game}}
    \end{floatrow}
\end{figure}

The decapsulation oracle $\mathcal{O}^\decap$ takes a ciphertext $c$ and returns the output of the $\decap$ routine using the secret key. The advantage $\epsilon_\monospace{IND-CCA}$ of an IND-CCA adversary $\mathcal{A}_\monospace{IND-CCA}$ is defined by

\begin{equation*}
    \epsilon_\monospace{IND-CCA} = \norm{
        P[\hat{b} = b] - \frac{1}{2}
    }
\end{equation*}

\subsection{Message authentication code}
A message authentication code $\mac$ is a collection of routines $(\sign, \verify)$ defined over some key space $\mathcal{K}$, some message space $\mathcal{M}$, and some tag space $\mathcal{T}$. The signing routine $\sign(k, m)$ takes the secret key $k \in \mathcal{K}$ and some message, and outputs a tag $t$. The verification routine $\verify(k, m, t)$ takes the triplet of secret key, message, and tag, and outputs $\monospace{1}$ if the message-tag pair is valid under the secret key, or $\monospace{0}$ otherwise.

@@

                \begin{algorithmic}[1]
                    \State $k^\ast \leftsample \mathcal{K}$
                    \State $(\hat{m}, \hat{t}) \leftsample \mathcal{A}^{\mathcal{O}_\sign}()$
                    \State \Return $
                        \llbrack \verify(k^\ast, \hat{m}, \hat{t}) 
                        \land (\hat{m}, \hat{t}) \not\in \mathcal{O}_\sign
                        \rrbrack
                    $
                \end{algorithmic}
            \end{algorithm}
        }{\caption{The \monospace{EUF-CMA} game}}
    \end{floatrow}
\end{figure}

The advantage $\epsilon_\monospace{EUF-CMA}$ of the existential forgery adversary is the probability that it wins the EUF-CMA game.

\subsection{Modular Fujisaki-Okamoto transformation}
The Fujisaki-Okamoto transformation (FOT) \cite{fujisaki1999secure} is a generic transformation that takes a $\pke$ with weaker security (such as \monospace{OW-CPA} or \monospace{IND-CPA}) and outputs a $\pke$ with stronger security. A later variation \cite{hofheinz2017modular} improved the original construction in \cite{fujisaki1999secure} by accounting for decryption failures, tightening security bounds, and providing a modular construction that first transforms \monospace{OW-CPA}/\monospace{IND-CPA} $\pke$ into \monospace{OW-PCVA} $\pke$ by providing ciphertext integrity through re-encryption (the \monospace{T} transformation), then transforming the \monospace{OW-PCVA} $\pke$ into an \monospace{IND-CCA} $\kem$ (the \monospace{U} transformation).

@@

Particularly relevant to our results are two variations of the \monospace{U} transformation: $\monospace{U}^\bot$ (KEM with explicit rejection) and $\monospace{U}^{\not\bot}$ (KEM with implicit rejection). If $\pke$ is \monospace{OW-PCVA} secure, then $\monospace{U}^\bot$ transforms $\pke$ into an \monospace{IND-CCA} secure $\kem^\bot$:

\begin{theorem}\label{thm:u-bot}
    For any \monospace{IND-CCA} adversary $\mathcal{A}_\kem$ against $\kem^\bot$ with advantage $\epsilon_\kem$ issuing at most $q_D$ decapsulation queries and at most $q_H$ hash queries, there exists an \monospace{OW-PCVA} adversary $\mathcal{A}_\pke$ against the underlying $\pke$ with advantage $\epsilon_\pke$ that makes at most $q_H$ queries to $\pco$ and $\cvo$ such that

@@

Similarly, if $\pke$ is \monospace{OW-PCA} secure, then $\monospace{U}^{\not\bot}$ transforms $\pke$ into an \monospace{IND-CCA} secure $\kem^{\not\bot}$

\begin{theorem}\label{thm:u-notbot}
    For any \monospace{IND-CCA} adversary $\mathcal{A}_\kem$ against $\kem^{\not\bot}$ with advantage $\epsilon_\kem$ issuing at most $q_D$ decapsulation queries and at most $q_H$ hash queries, there exists an \monospace{OW-CPA} adversary $\mathcal{A}_\pke$ against the underlying $\pke$ with advantage $\epsilon_\pke$ issuing at most $q_H$ queries to $\pco$ such that:

@@

The modularity of the \monospace{T} and \monospace{U} transformation allows us to tweak only the \monospace{T} transformation (see section \ref{sec:main-results}), obtain \monospace{OW-PCVA} security, then automatically get \monospace{IND-CCA} security for free. This means that we can directly apply our contribution to existing KEM's already using this modular transformation, such as ML-KEM \cite{key2023mechanism}, and obtain performance improvements while maintaining comparable levels of security (see section \ref{sec:mac-performance}).

@@

\begin{figure}[H]
    \center
    \begin{minipage}{0.29\textwidth}
        \begin{algorithm}[H]
            \caption{$\keygen_\etm$}\label{alg:etm-keygen}
            \begin{algorithmic}[1]
                \State $(\pk, \sk_\pke) \leftsample \keygen(1^\lambda)$
                \State $z \leftsample \mathcal{M}_\pke$
                \State $\sk \leftarrow (\sk_\pke, z)$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.32\textwidth}
        \begin{algorithm}[H]
            \caption{$\encap_\etm(\pk)$}\label{alg:etm-encap}
            \begin{algorithmic}[1]
                \State $m \leftsample \mathcal{M}_\pke$
                \State $k_\mac \leftarrow G(\pk, m)$
                \State $c_\pke \leftsample \encrypt(\pk, m)$
                \State $t \leftarrow \mac(k, c_\pke)$
                \State $K \leftarrow H(\pk, m, c_\pke)$
                \State $c \leftarrow (c_\pke, K)$
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.37\textwidth}
        \begin{algorithm}[H]
            \caption{$\decap_\etm(\sk, c)$}\label{alg:etm-decap}
            \begin{algorithmic}[1]
                \State $(c_\pke, t) \leftarrow c$
                \State $(\sk_\pke, z) \leftarrow \sk$
                \State $\hat{m} \leftarrow \decrypt(\sk_\pke, c_\pke)$
                \State $\hat{k}_\pk \leftarrow G(\pk, \hat{m})$
                \If{$\mac(\hat{k}, c_\pke) \neq t$}
                    \State \Return $H(\pk, z, c_\pke)$
                \EndIf
                \State \Return $H(\pk, \hat{m}, c_\pke)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{$\kem_\etm$ routines}\label{fig:etm-routines} 
\end{figure}

Here are a few design rationale:
\begin{enumerate}
    \item \emph{Deriving MAC key and session key}. We choose to include the public key $\pk$ when deriving the MAC key $k_\mac$ and the session key $K$ for similar reason stated in \cite{bos2018crystals}. If the MAC key is derived solely from the message, then an adversary can pre-compute a large lookup table mapping MAC key to the source plaintext that can applied to \emph{all sessions}. When the adversary intercepts a ciphertext $c = (c_\pke, t)$, it can brute-force all possible MAC keys to recover the plaintext. Since brute-forcing MAC key on a known ciphertext-tag pair is an offline search, given a sufficiently large lookup table and a sufficiently large quantum computer, this search might be feasible. On the other hand, if the MAC key is derived from both the public key and the plaintext, then the adversary will need to compute the large lookup table and run the key search \emph{per session}, which greatly increases the cost of the attack.
    \item \emph{Not hashing unauthenticated ciphertext $c_\pke$}: \cite{avanzi2019crystals}\cite{bos2018crystals} derives the session key from a hash of the ciphertext $K \leftarrow \monospace{KDF}(\overline{K} \Vert H(c_\pke))$ to \emph{``simplify implementation with non-incremental hash APIs''}. Since ``hashing the ciphertext separately'' vs ``deriving session key directly from ciphertext''  does not affect the security nor the performance (since the ciphertext will be hashed somewhere) of the scheme, we opt to not complicate the design with unnecessary hashes and leave the implementation considerations to concrete instantiations.
    \item \emph{Not adding key confirmation $\sigma \leftarrow \mac(\cdot, K)$}: within the security model of a KEM, there is no interaction between the KEM adversary and any other party (be it the challenger or any oracle) that involves the adversary sending a session key. In other words, there is no scenario in which tempering with any session key is meaningful, which means that the session key does not need protection.
\end{enumerate}

\subsection{If PKE is OW-PCA secure, then EtM is IND-CCA2 secure}

\begin{theorem}\label{thm:ow-pca-implies-kem-ind-cca2}
    For every \monospace{IND-CCA2} adversary $A$ against $\kem_\etm$ that makes at most $q_D$ decapsulation queries, there exists an \monospace{OW-PCA} adversary $B$ against the underlying $\pke$ such that

    \begin{equation*}
        \monospace{Adv}_\monospace{IND-CCA2}(A) \leq q_D \cdot \epsilon_\mac + 2 \cdot \monospace{Adv}_\monospace{OW-PCA}(B)
    \end{equation*}
\end{theorem}

@@

When $A$ receives the challenge ciphertext $(c^\ast, t^\ast)$ and unknown session key $K_b \leftsample \{K_0, K_1\}$, $A$ passes $c^\ast$ as the challenge ciphertext to $B$ as $B$'s challenge ciphertext. After $B$ returns the guess $\hat{m}$, $A$ computes the expected session key using $\hat{m}$ and $c^\ast$. If $\hat{m}$ is the correct decryption of $c^\ast$, then the expected session key should match the correct session key, in which case $A$ will correctly distinguish the true session key from the random session key. If $\hat{m}$ is not the correct decryption of $c^\ast$, then the expected session key will probably not match anything, so $A$ will always claim $K_b$ to be a random key. In other words, \emph{if $B$ wins the \monospace{OW-PCA} game, then $A$ wins the \monospace{IND-CCA2} game; if $B$ does not win, then $A$'s chance of winning is exactly $\frac{1}{2}$}. Therefore, the advantage of $A$ is at least that of $B$.

\section{Application to Kyber}
\subsection{``encrypt-then-MAC'' is not secure with Kyber}
CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} and ML-KEM \cite{key2023mechanism} are \monospace{IND-CCA2} secure key encapsulation mechanisms whose security depends on the hardness of the Module Learning with Error (MLWE) problem. For the construction of the \monospace{IND-CCA2} secure KEM, Kyber first constructs an \monospace{IND-CPA} public-key encryption scheme (which we will call \monospace{CPAPKE}), then applies the modular Fujisaki-Okamoto transformation \cite{hofheinz2017modular} to construct the key encapsulation mechanism (which we will call \monospace{CCAKEM}). Specifically, Kyber's round-3 submission uses the $U^{\not\bot}$ transformation, while ML-KEM uses the $U^{\not\bot}_m$ transformation. The routines of \monospace{CPAPKE} can be found in Algorithm 4, 5, 6 in \cite{avanzi2019crystals} and are largely identical between Kyber and ML-KEM.

\emph{\monospace{CPAPKE} is not \monospace{OW-PCA} secure}: What's important to know about \monospace{CPAPKE} is that it is not secure against plaintext-checking attack.

The main reason why a plaintext-checking attack works against \monospace{CPAPKE} is that for each known honest plaintext-ciphertext pair $(m, c)$, one can perturb the ciphertext and obtain a new ciphertext $c^\prime \leftarrow \monospace{addNoise}(c)$. Where the perturbation is small, the perturbed ciphertext can still correctly decrypt to the original plaintext $\decrypt_\monospace{CPAPKE}(\sk, c^\prime) = \decrypt_\monospace{CPAPKE}(\sk, c) = m$. Where the perturbation is large, the perturbed ciphertext will not decrypt to the original plaintext. The boundary between "small perturbation" and "large perturbation" depends on the values of the secret $\mathbf{s}$, so an adversary can probe the boundary in each of the $k \times n$ dimensions and learn the value of $\mathbf{s}$ coefficient-by-coefficient.

>>>>

% Custom commands
\newcommand{\pke}{\texttt{PKE}}
\newcommand{\keygen}{\texttt{Gen}}
\newcommand{\encrypt}{\texttt{Enc}}
\newcommand{\decrypt}{\texttt{Dec}}
\newcommand{\kem}{\texttt{KEM}}
\newcommand{\encap}{\texttt{Encap}}
\newcommand{\decap}{\texttt{Decap}}
\newcommand{\etm}{\texttt{EtM}}  % encrypt-then-mac
\newcommand{\mac}{\texttt{MAC}}
\newcommand{\sign}{\texttt{Sign}}
\newcommand{\verify}{\texttt{Verify}}
\newcommand{\pk}{\texttt{pk}}
\newcommand{\sk}{\texttt{sk}}
\newcommand{\pco}{\texttt{PCO}}
\newcommand{\cvo}{\texttt{CVO}}
\newcommand{\leftsample}{\stackrel{\$}{\leftarrow}}
\newcommand{\llbrack}{[\![}
\newcommand{\rrbrack}{]\!]}
\newcommand{\norm}[1]{\left\lvert #1 \right\rvert}
\newcommand{\fotplus}{\texttt{FOT+}}

@@

The problem of ciphertext integrity was solved in symmetric cryptography. Given a semantically secure symmetric cipher and an existentially unforgeable message authentication code, combining them using ``encrypt-then-mac'' provides authenticated encryption \cite{bellare2000authenticated}. We took inspiration from this strategy and applied a similar technique to provide ciphertext integrity for a public-key encryption scheme, which then translates to an \texttt{IND-CCA} secure $\kem$. Using a message authentication code for ciphertext integrity replaces the re-encryption step in decryption with the computation of a tag, which should offer significant performance improvements while maintaining comparable level of security.

The main challenge in applying ``encrypt-then-mac'' to public-key cryptography is the lack of a pre-shared $\mac$ key. We proposed to derive the shared $\mac$ key by hashing the plaintext message. We will prove in section \ref{sec:main-results} that, under the random oracle model, the $\mac$ key is securely hidden behind the hash function, and producing a valid pair of ciphertext and tag without full knowledge of the plaintext constitutes a forgery attack on the message authentication code. Thanks to the modular construction in \cite{hofheinz2017modular}, providing ciphertext integrity in the underlying encryption scheme gives us an \texttt{IND-CCA} secure $\kem$ for free.

@@

\subsubsection{Security}
We discuss the security of a $\pke$ using the sequence of games described in \cite{shoup2004sequences}. Specifically, we first define the $\texttt{OW-ATK}$ and the $\texttt{IND-CPA}$ game as they pertain to a public key encryption scheme. In later section we will define the $\texttt{IND-CCA}$ game as it pertains to a key encapsulation mechanism.

In the \texttt{OW-ATK} game, an adversary's goal is to recover the decryption of a randomly generated ciphertext.

@@

    \begin{algorithm}[H]
        \caption{
            The $\texttt{OW-ATK}$ game
        }\label{alg:pke-ow-atk-game}
        \begin{algorithmic}[1]
            \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
            \State $m^\ast \leftsample \mathcal{M}$
            \State $c^\ast \leftsample \encrypt(\pk, m)^\ast$
            \State $\hat{m} \leftsample \mathcal{A}^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
            \State \Return $\llbrack m^\ast = \hat{m} \rrbrack$
        \end{algorithmic}
    \end{algorithm}

    \caption{The \texttt{OW-ATK} game}
    \label{fig:pke-ow-atk-game}
\end{figure}

The adversary $\mathcal{A}$ with access to oracle(s) $\mathcal{O}_\texttt{ATK}$ wins the game if its guess $\hat{m}$ is equal to the challenge plaintext $m^\ast$. The \textit{advantage} $\epsilon_\texttt{OW-ATK}$ of an adversary in this game is the probability that it wins the game.

The choice of oracle(s) $\mathcal{O}_\texttt{ATK}$ depends on the choice of $\texttt{ATK}$. Specifically:

\begin{equation*}
    \mathcal{O}_\texttt{ATK} = \begin{cases}
        - & \texttt{ATK} = \texttt{CPA} \\
        \texttt{PCO} & \texttt{ATK} = \texttt{PCA} \\
        \texttt{CVO} & \texttt{ATK} = \texttt{VA} \\
        \texttt{PCO, CVO} & \texttt{ATK} = \texttt{PCVA}
    \end{cases}
\end{equation*}

Where the definitions of plaintext-checking oracle $\texttt{PCO}$ and the ciphertext-validation oracle $\texttt{CVO}$ are inherited from \cite{hofheinz2017modular}

@@

In the \texttt{IND-CPA} game (algorithm \ref{alg:ind-cpa-game}), an adversary's goal is to distinguish the encryption of one message from the encryption of another message. Given the public key, the adversary outputs two adversarially chosen messages and obtains the encryption of a random choice between these two messages. The adversary wins the \texttt{IND-CPA} game if it correctly identifies which message the encryption is obtained from.

\begin{figure}[H]
    \begin{floatrow}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{The $\texttt{IND-CPA}$ game}\label{alg:ind-cpa-game}

                \begin{algorithmic}[1]
                    \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                    \State $(m_0, m_1) \leftsample \mathcal{A}(a^\lambda, \pk)$
                    \State $b \leftsample \{0, 1\}$
                    \State $c^\ast \leftsample \encrypt(\pk, m_b)$
                    \State $\hat{b} \leftsample \mathcal{A}(1^\lambda, \pk, c^\ast)$
                \State \Return $\llbrack b = \hat{b} \rrbrack$
            \end{algorithmic}
        \end{algorithm}
        }{\caption{The \texttt{IND-CPA} game}}
    \end{floatrow}
\end{figure}

The \textit{advantage} $\epsilon_\texttt{IND-CPA}$ of an \texttt{IND-CPA} adversary $A$ is defined by

$$
    \texttt{Adv}_\texttt{IND-CPA}(A) = \norm{
        P[\hat{b} = b] - \frac{1}{2}
    }
$$

@@

\begin{figure}[H]
    \begin{floatrow}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{
                    \texttt{IND-CCA} game for \kem
                }\label{alg:kem-ind-cca-game}

                \begin{algorithmic}[1]
                    \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                    \State $(c^\ast, k_0) \leftsample \encap(\pk)$
                    \State $k_1 \leftsample \mathcal{K}$
                    \State $b \leftsample \{0, 1\}$
                    \State $\hat{b} \leftsample \mathcal{A}_\texttt{IND-CCA}^{\mathcal{O}_\decap}(
                        1^\lambda, \pk, c^\ast, k_b
                    )$
                    \State \Return $\llbrack \hat{b} = b \rrbrack$
                \end{algorithmic}
            \end{algorithm}
        }{\caption{The \texttt{KEM-IND-CCA2} game}}
    \end{floatrow}
\end{figure}

The decapsulation oracle $\mathcal{O}^\decap$ takes a ciphertext $c$ and returns the output of the $\decap$ routine using the secret key. The advantage $\epsilon_\texttt{IND-CCA}$ of an IND-CCA adversary $\mathcal{A}_\texttt{IND-CCA}$ is defined by

\begin{equation*}
    \epsilon_\texttt{IND-CCA} = \norm{
        P[\hat{b} = b] - \frac{1}{2}
    }
\end{equation*}

\subsection{Message authentication code}
A message authentication code $\mac$ is a collection of routines $(\sign, \verify)$ defined over some key space $\mathcal{K}$, some message space $\mathcal{M}$, and some tag space $\mathcal{T}$. The signing routine $\sign(k, m)$ takes the secret key $k \in \mathcal{K}$ and some message, and outputs a tag $t$. The verification routine $\verify(k, m, t)$ takes the triplet of secret key, message, and tag, and outputs $\texttt{1}$ if the message-tag pair is valid under the secret key, or $\texttt{0}$ otherwise.

@@

                \begin{algorithmic}[1]
                    \State $k^\ast \leftsample \mathcal{K}$
                    \State $(\hat{m}, \hat{t}) \leftsample \mathcal{A}^{\mathcal{O}_\sign}()$
                    \State \Return $
                        \llbrack \verify(k^\ast, \hat{m}, \hat{t}) 
                        \land (\hat{m}, \hat{t}) \not\in \mathcal{O}_\sign
                        \rrbrack
                    $
                \end{algorithmic}
            \end{algorithm}
        }{\caption{The \texttt{EUF-CMA} game}}
    \end{floatrow}
\end{figure}

The advantage $\epsilon_\texttt{EUF-CMA}$ of the existential forgery adversary is the probability that it wins the EUF-CMA game.

\subsection{Modular Fujisaki-Okamoto transformation}
The Fujisaki-Okamoto transformation (FOT) \cite{fujisaki1999secure} is a generic transformation that takes a $\pke$ with weaker security (such as \texttt{OW-CPA} or \texttt{IND-CPA}) and outputs a $\pke$ with stronger security. A later variation \cite{hofheinz2017modular} improved the original construction in \cite{fujisaki1999secure} by accounting for decryption failures, tightening security bounds, and providing a modular construction that first transforms \texttt{OW-CPA}/\texttt{IND-CPA} $\pke$ into \texttt{OW-PCVA} $\pke$ by providing ciphertext integrity through re-encryption (the \texttt{T} transformation), then transforming the \texttt{OW-PCVA} $\pke$ into an \texttt{IND-CCA} $\kem$ (the \texttt{U} transformation).

@@

Particularly relevant to our results are two variations of the \texttt{U} transformation: $\texttt{U}^\bot$ (KEM with explicit rejection) and $\texttt{U}^{\not\bot}$ (KEM with implicit rejection). If $\pke$ is \texttt{OW-PCVA} secure, then $\texttt{U}^\bot$ transforms $\pke$ into an \texttt{IND-CCA} secure $\kem^\bot$:

\begin{theorem}\label{thm:u-bot}
    For any \texttt{IND-CCA} adversary $\mathcal{A}_\kem$ against $\kem^\bot$ with advantage $\epsilon_\kem$ issuing at most $q_D$ decapsulation queries and at most $q_H$ hash queries, there exists an \texttt{OW-PCVA} adversary $\mathcal{A}_\pke$ against the underlying $\pke$ with advantage $\epsilon_\pke$ that makes at most $q_H$ queries to $\pco$ and $\cvo$ such that

@@

Similarly, if $\pke$ is \texttt{OW-PCA} secure, then $\texttt{U}^{\not\bot}$ transforms $\pke$ into an \texttt{IND-CCA} secure $\kem^{\not\bot}$

\begin{theorem}\label{thm:u-notbot}
    For any \texttt{IND-CCA} adversary $\mathcal{A}_\kem$ against $\kem^{\not\bot}$ with advantage $\epsilon_\kem$ issuing at most $q_D$ decapsulation queries and at most $q_H$ hash queries, there exists an \texttt{OW-CPA} adversary $\mathcal{A}_\pke$ against the underlying $\pke$ with advantage $\epsilon_\pke$ issuing at most $q_H$ queries to $\pco$ such that:

@@

The modularity of the \texttt{T} and \texttt{U} transformation allows us to tweak only the \texttt{T} transformation (see section \ref{sec:main-results}), obtain \texttt{OW-PCVA} security, then automatically get \texttt{IND-CCA} security for free. This means that we can directly apply our contribution to existing KEM's already using this modular transformation, such as ML-KEM \cite{key2023mechanism}, and obtain performance improvements while maintaining comparable levels of security (see section \ref{sec:mac-performance}).

@@

\begin{figure}[H]
    \center
    \begin{algorithm}[H]
        \caption{$\keygen_\etm$}\label{alg:etm-keygen}
        \begin{algorithmic}[1]
            \State $(\pk, \sk_\pke) \leftsample \keygen(1^\lambda)$
            \State $z \leftsample \mathcal{M}_\pke$
            \State $\sk \leftarrow (\sk_\pke, z)$
            \State \Return $(\pk, \sk)$
        \end{algorithmic}
    \end{algorithm}
    \begin{algorithm}[H]
        \caption{$\encap_\etm(\pk)$}\label{alg:etm-encap}
        \begin{algorithmic}[1]
            \State $m \leftsample \mathcal{M}_\pke$
            \State $k \leftarrow G(m)$
            \State $c_\pke \leftsample \encrypt(\pk, m)$
            \State $t \leftarrow \mac(k, c_\pke)$
            \State $K \leftarrow H(m, c_\pke)$
            \State $c \leftarrow (c_\pke, K)$
            \State \Return $(c, K)$
        \end{algorithmic}
    \end{algorithm}
    \begin{algorithm}[H]
        \caption{$\decap_\etm(\sk, c)$}\label{alg:etm-decap}
        \begin{algorithmic}[1]
            \State $(c_\pke, t) \leftarrow c$
            \State $(\sk_\pke, z) \leftarrow \sk$
            \State $\hat{m} \leftarrow \decrypt(\sk_\pke, c_\pke)$
            \State $\hat{k} \leftarrow G(\hat{m})$
            \If{$\mac(\hat{k}, c_\pke) \neq t$}
                \State \Return $H(z, c_\pke)$
            \EndIf
            \State \Return $H(\hat{m}, c_\pke)$
        \end{algorithmic}
    \end{algorithm}
    \caption{$\kem_\etm$ routines}\label{fig:etm-routines} 
\end{figure}

\begin{theorem}\label{thm:ow-pca-implies-kem-ind-cca2}
    For every \texttt{IND-CCA2} adversary $A$ against $\kem_\etm$ that makes at most $q_D$ decapsulation queries, there exists an \texttt{OW-PCA} adversary $B$ who makes $q_P$ plaintext-checking queries against the underlying $\pke$ such that

    \begin{equation*}
        \texttt{Adv}_\texttt{IND-CCA2}(A) \leq q_D \cdot \epsilon_\mac + 2 \cdot \texttt{Adv}_\texttt{OW-PCA}(B)
    \end{equation*}
\end{theorem}

@@

When $A$ receives the challenge ciphertext $(c^\ast, t^\ast)$ and unknown session key $K_b \leftsample \{K_0, K_1\}$, $A$ passes $c^\ast$ as the challenge ciphertext to $B$ as $B$'s challenge ciphertext. After $B$ returns the guess $\hat{m}$, $A$ computes the expected session key using $\hat{m}$ and $c^\ast$. If $\hat{m}$ is the correct decryption of $c^\ast$, then the expected session key should match the correct session key, in which case $A$ will correctly distinguish the true session key from the random session key. If $\hat{m}$ is not the correct decryption of $c^\ast$, then the expected session key will probably not match anything, so $A$ will always claim $K_b$ to be a random key. In other words, \emph{if $B$ wins the \texttt{OW-PCA} game, then $A$ wins the \texttt{IND-CCA2} game; if $B$ does not win, then $A$'s chance of winning is exactly $\frac{1}{2}$}. Therefore, the advantage of $A$ is at least that of $B$.

\section{Application to Kyber}

\subsection{Kyber with AE}
CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} and ML-KEM \cite{key2023mechanism} are \texttt{IND-CCA2} secure key encapsulation mechanisms whose security depends on the hardness of the Module Learning with Error (MLWE) problem. For the construction of the \texttt{IND-CCA2} secure KEM, Kyber first constructs an \texttt{IND-CPA} public-key encryption scheme (which we will call \texttt{CPAPKE}), then applies the modular Fujisaki-Okamoto transformation \cite{hofheinz2017modular} to construct the key encapsulation mechanism (which we will call \texttt{CCAKEM}). Specifically, Kyber's round-3 submission uses the $U^{\not\bot}$ transformation, while ML-KEM uses the $U^{\not\bot}_m$ transformation. The routines of \texttt{CPAPKE} can be found in Algorithm 4, 5, 6 in \cite{avanzi2019crystals} and are largely identical between Kyber and ML-KEM.

The routines of \texttt{CCAKEM} can be found in Algorithm 7, 8, 9 in \cite{avanzi2019crystals}. We modify Algorithms 8 and 9 using authenticated encryption (AE)  mode as follows where Algorithm 7 is unchanged.

\begin{algorithm}[H]
    \caption{\texttt{Kyber.CCAKEM.KeyGen()}}
    \begin{algorithmic}[1]
        \State $z \leftsample \mathcal{B}^{32}$
            \Comment{Randomly sample 32 bytes (256 bits)}
        \State $(\pk, \sk^\prime) \leftsample \texttt{Kyber.CPAPKE.KeyGen()}$
        \State $\sk = (\sk^\prime, \pk, H(\pk), z)$
            \Comment{H is instantiated with SHA3-256}
        \State \Return $(\pk, \sk)$
    \end{algorithmic}
\end{algorithm}
